---
title: "Equations of Straight Line Graphs - Part 2"
description: |
  This lesson follows on from ‘Using R to draw the equation of a straight line graph - Part 1’ and demonstrates how to use R to plot straight line graphs (y=mx+c) with different intercepts and negative gradients followed by questions on the observed differences.
# author: []
date: 2022-02-25
output:
  distill::distill_article:
    self_contained: false
---

`r xfun::embed_file(here::here("_math", "2022-02-25-straightline2post", "equation_straight_line2.Rmd"), text = "Download .Rmd here")`

`r xfun::embed_file(here::here("_math", "2022-02-25-straightline2post", "equation_straight_line2.docx"), text = "Download word document here")`

`r xfun::embed_file(here::here("_math", "2022-02-25-straightline2post", "equation_straight_line2.pdf"), text = "Download pdf here")`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## KS3: Plotting and finding the equations of straight line graphs (y=mx+c) - Part 2

## Lesson objectives
Plotting and finding the equations of straight line graphs y=mx+c

## Success criteria
* Plot a straight line graph y = mx+c
* Plot straight line graphs y = mx+c with different intercepts
* Plot a straight line graph y = mx+c with a negative gradient
* Find the equations of straight line graphs

## Keywords
* slope
* gradient
* intercept
* x-axis
* y-axis

## Worked Example 1
This is a worked example for you to follow.

We will show you how to plot the line graph for y=2x + 5.

```{r chunk1}
x <- seq(from=-4, to=4, by=1) # sequence the x-axis from -4 to 4
y <- 2*x+5
mydata <- tibble (x,y)
ggplot(mydata) +
  aes(x,y) +
  geom_point () +
  geom_line (col='red')+
  geom_vline (xintercept = 0, col='black')+
  geom_hline (yintercept = 0, col='black')

```


## Worked Example 2
We will now see what happens when you change the intercept c to a negative value.

1. y=3x+5
2. y=3x-5

Run the code in chunk2 by clicking on the little arrow on the right of the code chunk.


```{r chunk2}
x <- seq(-4, 4) # sequence from -4 to 4
y1 <- 3*x+5
y2 <- 3*x-5
mydata <- tibble (x,y1,y2,y)
ggplot(mydata) +
  geom_line (aes(x=x, y=y1), col='red')+
  geom_line (aes(x=x, y=y2), col='blue')+
  geom_vline (xintercept = 0, col='black')+
  geom_hline (yintercept = 0, col='black')

```
Question: What is the same and what is different about these lines?

Answer:



Knit your document and check the output.


## Activity 1

Draw these lines on a graph.  Use the R code from chunk2 to help you.  Remember to update the code with these new lines.

1. y=x+10
2. y=x-10

```{r chunk3}


```
Question: What is the same and what is different about these lines?

Answer:




## Activity 2
Now we will investigate m (the gradient). Draw these lines on a graph. Use R code from chunk2 to help you. 

Run the code and knit the document.

1. y=-2x+5
2. y= 2x+5

```{r chunk4}


```
Question: What is the same and what is different about these lines?

Answer:








## Activity 3 Answer the following questions

```{r chunk5 questions, echo=FALSE}
cat ("Q1 In the equation y=mx+c, what happens when you change c?")
cat ("Q2 What happens when you change m?")
```

Write your answers here\
Q1:

Q2:


## Activity 4: Work out the equation from a line graph

Write down the equations of the following four lines on the graph below.

Knit the document to get a good view of the graph.

```{r chunk6, echo=FALSE}
x <- seq(-10, 10)
y <- x
ggplot() +
  aes(x,y)+
  geom_blank()+
  geom_abline(slope=1, intercept=0, col='red')+
  geom_abline(slope=-1, intercept=2.5, col='cyan')+
  geom_abline(slope=2, intercept=10, col='blue')+
  geom_abline(slope=0, intercept=5, col='orange')+
  geom_vline (xintercept=0, col='black')+
  geom_hline (yintercept=0, col='black')

```

Write your answers here

Cyan line:

Red line:

Blue line:

Orange line:


KNIT YOUR DOCUMENT for the final time.  This will be the version that your teacher will mark.

## THE END

