---
title: "With thanks to:"
# description: |
#   A new article created using the Distill format.
# author:
#   - name: Nora Jones 
#     url: https://example.com/norajones
#     affiliation: Spacely Sprockets
#     affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
#remotes::install_github("EllaKaye/distilltools")
#library(distilltools)
```


<div style="text-align:left">
<a class="contributor_org" href="https://nhsrcommunity.com/" target="_blank">NHS-R Community (Professor Mohammed A Mohammed)</a>
</div>


<div style="text-align:left">
<a class="contributor_org" href="https://greenoak.bham.sch.uk/" target="_blank">Green Oak Academy</a>
<br><br>
<ul >
<li class="contributor_ppl" style="margin-left: 25px"> Maulana Maseehullah Patel (Executive Head Teacher) </li>
<li class="contributor_ppl" style="margin-left: 25px"> Dr Razia Ghani (Head Teacher) Green Oak Academy </li>
</ul>
</div>

<div style="text-align:left">
<a class="contributor_org" href="https://www.r-consortium.org/" target="_blank">R Consortium</a>
</div>

<div style="text-align:left">
<a class="contributor_org" href="https://www.mango-solutions.com/" target="_blank">Mango solutions</a>
</div>


<div style="text-align:left" class="contributor_ppl" style="padding-bottom: 0px;">
Ms A Nawaz (for designing R Girls logo) <br><br>
<a href="https://jennysloane.netlify.app/" target="_blank">Jenny Sloane</a> (for developing R Girls website) <br><br>
<a href="https://batool-almarzouq.netlify.app/" target="_blank">Batool Almarzouq</a> (for developing the interactive learnr lessons) <br><br>
Iman Al-Hasani <br><br>
Nasrin Fathollahzadeh Attar <br><br>
Mrs M Nawaz <br><br>


*This is a world wide community project for girls and if you wish to contribute please get in touch via slack, twitter or email*

</div>
